---
title: |
  | \vspace{-4em}MAT5314 Project 3: Queueing Theory
author: 
  - Teng Li(7373086)
  - Shiya Gao(300381032) 
  - Chuhan Yue(300376046)
  - Yang Lyu(8701121)
output: 
  pdf_document: 
    keep_tex: true
    includes:
      in_header: columns.tex
fontsize: 12pt
header-includes: 
  - \renewcommand{\and}{\\}
  - \usepackage{float}
  - \floatplacement{figure}{H}
bibliography: References.bib
link-citations: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
library(plotly)
library(tidyverse)
library(kableExtra)
library(htmlwidgets)
library(webshot)
```

### Introduction

It is common that a passenger gets screened at an airfield for security and management purpose. This screening process involves some major steps such as scanning the boarding pass at entrance, serving the passenger, and exiting the service desk. This process can be described by queuing theory, in which a statistical model is built to analyze the waiting time and service processes. It allows to examine the relationship between inter-arrivals, service times and the capacity of the system demonstrated by calculating the number of servers needed to achieve certain quality of service level. In the end, one can gain an overall picture of the efficiency and dynamics of the screening process.

We were given four data sets containing the details of the screening process at the four major airfields of the nation. Our goal was to explain the data sets through data definitions and explanations, to explore the data sets by visualizing and analyzing their variables and possible relationships, and finally to perform the queuing analysis to predict the expected wait time at each airfield.

### Data Definition

```{r}
data_1<-read.csv("BASA_AUC_2028_912.csv", header = TRUE)
data_2<-read.csv("dat_F_sub.csv")
data_3<-read.csv("dat_P_sub_c.csv")
data_4<-read.csv("years20262030.csv")
```

The four data sets had common variables. The AUC data set was the biggest one containing most of the common variables, hence we used it as the main data set. We compared the variables between AUC and Passenger data and noticed that Passenger had unique variables with suffix "Flag", which we thought was not useful for our analysis. Other unique variables in Passenger were presented in other tables as well, hence omitted. 
```{r}
setdiff(colnames(data_3), colnames(data_1))
```

Similarly, the Flight data had variables that weren't really relevant or useful for our analysis, hence we also omitted them.
```{r}
setdiff(colnames(data_2), colnames(data_1))
```

AUC and Y2026 had the same variables, except that Y2026 has airfields other than Auckland.

We gave a proper data definition in the following Table(). 

```{r}
data_1<-read.csv("BASA_AUC_2028_912.csv", header = TRUE)%>%select(-X)
DataDict<-data.frame(
  Variables=colnames(data_1),
  Example=sapply(data_1, function(x) paste(as.character(head(unique(x),2)), collapse = ", ")),
  Number.Unique=sapply(data_1, function(x) length(unique(x))),
  PctMissing=sapply(data_1, function(x) paste0(round(sum(is.na(x))/length(x), 4)*100,"%" ) ),
  Comment=c("Airfield from which the passenger departed: Auckland [AUC], Chebucto [CWL], Queenston [QUE], SaintFrancois [SAF]",
            "the date and time at which passengers exited the main queue S2, recorded to the nearest minute;",
            "The waiting time of passenger in queue from S1 to S2, rounded up to the minute",
            "The number of security agents at S1 when passenger was scaned",
            "The number of security agents at S2 when passenger was scaned",
            "The average number of security agents at S1 and S2",
            "scheduled departure time",
            "actual departure time" ,
            "city of destination",
            "country of destination",
            "order",
            "ID for each passenger exited the main queue",
            "Departure_Date",
            "Departure_Time (recored in seconds: h*3600+m*60+s)",
            "departure time of day",
            "departure day is weekend or weeakday",
            "departure day in a week",
            "month of departure",
            "season of departure",
            "year of departure"
            ), row.names = NULL
)

DataDict%>%kable(booktabs = TRUE,caption = "AUC Variable Definition")%>%
  kable_styling(font_size=10, latex_options=c("striped", "scale_down","hold_position"))%>%
  column_spec(2, width = "8em")%>%
  row_spec(0,bold=TRUE)
```

```{r}
data_2<-read.csv("dat_F_sub.csv")%>%select(-X)
DataDict<-data.frame(
  Variables=colnames(data_2),
  Example=sapply(data_2, function(x) paste(as.character(head(unique(x),5)), collapse = ", ")),
  Number.Unique=sapply(data_2, function(x) length(unique(x))),
  PctMissing=sapply(data_2, function(x) paste0(round(sum(is.na(x))/length(x), 4)*100,"%" ) ),
  Comment=c("Airfield: Auckland [AUC], Chebucto [CWL], Queenston [QUE], SaintFrancois [SAF]",
            "Flight ID",
            "scheduled time of departure",
            "the actual departure time of each passenger’s flight" ,
            "departure time in one day",
            "departrure day is weekend or weeakday",
            "departure day in a week",
            "month",
            "season",
            "year",
            "total number of passengers",
            "actual number of passengers",
            "mimimum waiting time",
            "mean waiting time",
            "median waiting time",
            "maximum waiting time",
            "mean length of waiting time",
            "mean_City_Flag",
            "mode_BFO_Dest_City",
            "sum of the mode in the city",
            "number of destinated city",
            "mode_BFO_Dest_Country_Code",
            "sum of the mode in the country",
            "number of destinated country",
            "delay second"
            ), row.names = NULL
)

DataDict%>%kable(booktabs = TRUE,caption = "dat\\_F Variable Definition")%>%
  kable_styling(font_size=10, latex_options=c("striped","scale_down","hold_position"))%>%
  column_spec(4, width = "8em")%>%
  row_spec(0,bold=TRUE)
```

```{r}
data_3<-read.csv("dat_P_sub_c.csv")%>%select(-Pass_ID)
DataDict<-data.frame(
  Variables=colnames(data_3),
  Example=sapply(data_3, function(x) paste(as.character(head(unique(x),5)), collapse = ", ")),
  Number.Unique=sapply(data_3, function(x) length(unique(x))),
  PctMissing=sapply(data_3, function(x) paste0(round(sum(is.na(x))/length(x), 4)*100,"%" ) ), row.names = NULL
)

DataDict%>%kable(booktabs = TRUE,caption = "dat\\_P Variable Definition")%>%
  kable_styling(font_size=10, latex_options=c("striped","scale_down","hold_position"))%>%
  column_spec(4, width = "8em")%>%
  row_spec(0,bold=TRUE)
```


```{r}
result <- AUC %>%
  group_by(Year, Season, Time_of_Day, Day_of_Week) %>%  # 按照年、季度和时间段分组
  summarise(
   Meantime= mean(Wait_Time),# 计算均值或其他汇总统计量
   Meanservers=mean(C0),
   pro_ser1=sum(C0== 1)/n(),
   pro_ser2=sum(C0== 2)/n(),
   pro_ser3=sum(C0== 3)/n(),
   ave_ser=pro_ser1*1+pro_ser2*2+pro_ser2*3,
   TotalCount = n(),# 统计每个组的数据数量
    .groups = "keep"
  )

head(result)
```

```{r}
#average arrival rate, servers, wait time, service rate
lambda <-mean(data_vector)  # 平均到达率（每小时到达的人数）
# 假设 observed 是你的观察值向量， expected 是期望值向量
chisq.test(observed, p = lambda)

mu <- 6     # 平均服务率（每小时服务的人数）
simulation_time <- 1000  # 模拟时间（小时）

# 初始化模拟数据
clock <- 0
queue <- 0
departure <- Inf

# 统计信息
total_customers <- 0
total_waiting_time <- 0

# 开始模拟
while (clock < simulation_time) {
  if (clock < departure) {
    interarrival_time <- rexp(1, 1 / lambda)
    clock <- clock + interarrival_time
    total_customers <- total_customers + 1
    if (queue == 0) {
      service_time <- rexp(1, 1 / mu)
      departure <- clock + service_time
    } else {
      queue <- queue + 1
    }
  } else {
    queue <- queue - 1
    total_waiting_time <- total_waiting_time + (departure - clock)
    service_time <- rexp(1, 1 / mu)
    clock <- departure
    departure <- clock + service_time
  }
}

# 计算模拟结果
average_waiting_time <- total_waiting_time / total_customers
average_queue_length <- total_waiting_time / simulation_time

cat("模拟结果:\n")
cat("平均等待时间: ", average_waiting_time, "小时\n")
cat("平均队列长度: ", average_queue_length, "\n")
```

```{r}
data_4<-read.csv("years20262030.csv")
```


### Data exploration

### Queuing Analysis

### References



